---
title: "Infodemiology Analysis of Alpha-Gal Syndrome Awareness in Australia"
author: "Alexander W. Gofton"
date: today
format:
  gfm:
    toc: true
    toc-depth: 3
  html:
    page-layout: full
    toc: true
    toc-depth: 3
    code-fold: true
    self-contained: true
    fig-width: 10
    fig-height: 7
    theme:
      light: cosmo
      dark: darkly
execute:
  warning: false
  message: false
  cache: false
knitr:
  opts_chunk:
    results: hold
---

```{r setup}
library(tidyverse)
library(patchwork)
library(mgcv)
library(forecast)
library(tseries)
library(lmtest)
library(sandwich)
library(segmented)
library(zoo)
library(scales)
library(vars)

home_dir <- "/Users/gof005/Library/CloudStorage/OneDrive-CSIRO/OneDrive - Docs/01_Projects/alpha_gal/03_aGal_infoepi"
data_dir <- file.path(home_dir, "data")
processed_dir <- file.path(home_dir, "processed_data")
media_dir <- file.path(home_dir, "mediacloud")
fig_dir <- file.path(home_dir, "figures")

search_terms <- c(
  "alpha_gal",
  "alpha-gal_syndrome",
  "food_allergy",
  "mammalian_meat_allergy",
  "paralysis_tick",
  "red_meat_allergy",
  "meat_allergy",
  "tick_allergy"
)

# Nice labels for plots
term_labels <- c(
  "alpha_gal" = "Alpha-gal",
  "alpha-gal_syndrome" = "Alpha-gal syndrome",
  "food_allergy" = "Food allergy",
  "mammalian_meat_allergy" = "Mammalian meat allergy",
  "paralysis_tick" = "Paralysis tick",
  "red_meat_allergy" = "Red meat allergy",
  "meat_allergy" = "Meat allergy",
  "tick_allergy" = "Tick allergy"
)

# Download dates
download_dates <- sprintf("2026-02-%02d", 13:28)
```

```{r}
# Primary CSIRO Colours
MiddayBlue <- "#00A9CE"
MidnightBlue <- "#00313C"
Steel <- "#757579"
Mist <- "#DADBDC"
Black <- "#000000"

# Secondary CSIRO Colours
Blueberry <- "#1F22AA"
OceanBlue <- "#014B87"
Plum <- "#6D2077"
Teal <- "#017377"
Mint <- "#007A53"

# Accent CSIRO Colours
Fuschia <- "#DF1A95"
Orange <- "#E87721"
Gold <- "#FFB81C"
Lavender <- "#9FAEE5"
LightMint <- "#71CC98"
Forest <- "#78BE21"
LightTeal <- "#2DCBD3"

```

# 1. Data Loading and Averaging

## 1.1 Load Google Trends Data

We collected Google Trends RSV data for 8 search terms across 16 consecutive days (Feb 13--28, 2026), following the multi-day sampling protocol recommended by Romeiser et al. (2024) and Holzl et al. (2024) to account for Google Trends' stochastic sampling variability.

```{r load-trends}
# Load all daily downloads for all search terms
all_downloads <- list()

for (date_str in download_dates) {
  for (term in search_terms) {
    file_path <- file.path(
      data_dir,
      paste0("google_trends_monthly_au_", term, "_", date_str, ".csv")
    )

    if (file.exists(file_path)) {
      df <- read_csv(file_path, show_col_types = FALSE)
      colnames(df)[2] <- "rsv"

      # Parse date column
      if (is.character(df$date)) {
        if (grepl("^\\d{1,2}/\\d{1,2}/\\d{4}$", df$date[1])) {
          df$date <- as.Date(df$date, format = "%d/%m/%Y")
        } else {
          df$date <- as.Date(df$date)
        }
      }

      df$search_term <- term
      df$download_date <- date_str
      all_downloads[[paste(term, date_str)]] <- df
    }
  }
}

trends_raw <- bind_rows(all_downloads)

cat("Loaded", n_distinct(trends_raw$download_date), "daily downloads for",
    n_distinct(trends_raw$search_term), "search terms\n")
cat("Date range:", as.character(min(trends_raw$date)), "to",
    as.character(max(trends_raw$date)), "\n")
cat("Total observations:", nrow(trends_raw), "\n")
```

## 1.2 Average Across Downloads and Assess Sampling Variability

```{r average-trends}
# Average RSV across all 16 downloads per term per month
trends_avg <- trends_raw |>
  group_by(date, search_term) |>
  summarise(
    mean_rsv = mean(rsv, na.rm = TRUE),
    sd_rsv = sd(rsv, na.rm = TRUE),
    cv_rsv = ifelse(mean_rsv > 0, sd_rsv / mean_rsv * 100, NA),
    se_rsv = sd_rsv / sqrt(n()),
    lower_ci = mean_rsv - 1.96 * se_rsv,
    upper_ci = mean_rsv + 1.96 * se_rsv,
    n_downloads = n(),
    .groups = "drop"
  ) |>
  mutate(
    lower_ci = pmax(lower_ci, 0),
    search_term_label = term_labels[search_term]
  )

# Sampling variability summary
cv_summary <- trends_avg |>
  filter(mean_rsv > 0) |>
  group_by(search_term) |>
  summarise(
    median_cv = median(cv_rsv, na.rm = TRUE),
    mean_cv = mean(cv_rsv, na.rm = TRUE),
    max_cv = max(cv_rsv, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(median_cv)

cat("Sampling variability (CV%) across 16 downloads:\n")
print(cv_summary, n = Inf)

# Flag terms with high sampling variability
unreliable_terms <- cv_summary |>
  filter(median_cv > 100) |>
  pull(search_term)

reliable_terms <- setdiff(search_terms, unreliable_terms)

cat("\nTerms with median CV > 100% (unreliable month-to-month precision):\n")
cat(" ", paste(term_labels[unreliable_terms], collapse = ", "), "\n")
cat("These terms have low absolute search volume, causing Google Trends'",
    "stochastic sampling\nto produce highly variable RSV estimates.",
    "Results for these terms should be interpreted cautiously.\n")
```

## 1.3 Load MediaCloud Data

```{r load-media}
# Daily media article counts
media_counts <- read_csv(
  file.path(media_dir, "mc-onlinenews-mediacloud-20260213000613-counts.csv"),
  show_col_types = FALSE
) |>
  mutate(date = as.Date(date))

# Aggregate to monthly
media_monthly <- media_counts |>
  mutate(month_date = floor_date(date, "month")) |>
  group_by(month_date) |>
  summarise(
    articles = sum(count),
    total_articles = sum(total_count),
    .groups = "drop"
  ) |>
  rename(date = month_date)

# Media content (individual articles)
media_content <- read_csv(
  file.path(media_dir, "mc-onlinenews-mediacloud-20260213000622-content.csv"),
  show_col_types = FALSE
) |>
  mutate(publish_date = as.Date(publish_date))

# Top sources
media_sources <- read_csv(
  file.path(media_dir, "mc-onlinenews-mediacloud-20260213000707-top-sources.csv"),
  show_col_types = FALSE
)

cat("Media data: ", nrow(media_counts), " daily records,",
    sum(media_counts$count), "total articles\n")
cat("Date range:", as.character(min(media_counts$date)), "to",
    as.character(max(media_counts$date)), "\n")
cat("Individual articles in content data:", nrow(media_content), "\n")
```

# 2. Descriptive Trends in Search Interest

## 2.1 Time Series of All Search Terms

```{r fig-trends-faceted, fig.width=12, fig.height=10}
#| fig-cap: "Average Google Trends RSV (0-100) for AGS/MMA-related search terms in Australia, 2014-2026. Ribbons show 95% confidence intervals from 16 daily data downloads."

ggplot(trends_avg, aes(x = date, y = mean_rsv)) +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci),
              fill = "#2c7fb8", alpha = 0.2) +
  geom_line(colour = "#2c7fb8", linewidth = 0.6) +
  facet_wrap(~ search_term_label, scales = "free_y", ncol = 2) +
  labs(
    title = "Google Trends Search Interest for AGS/MMA-Related Terms in Australia",
    subtitle = paste0("Monthly RSV averaged across ", max(trends_avg$n_downloads),
                      " daily downloads (Feb 13-28, 2026)"),
    x = NULL, y = "Mean Relative Search Volume (RSV)",
    caption = "Ribbon = 95% CI | Source: Google Trends (Health category, Australia)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    strip.text = element_text(face = "bold", size = 10),
    panel.grid.minor = element_blank(),
    panel.spacing = unit(1, "lines"),
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(colour = "grey40"),
    plot.caption = element_text(hjust = 0, face = "italic", size = 8)
  ) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y")

ggsave(file.path(fig_dir, "fig1_trends_faceted.png"),
       width = 30, height = 22, units = "cm", dpi = 300)
```

## 2.2 AGS-Specific Terms Combined

```{r fig-ags-terms, fig.width=10, fig.height=6}
#| fig-cap: "Comparison of AGS-specific search terms showing the dual Australian/international terminology landscape."

ags_terms <- c("alpha_gal", "alpha-gal_syndrome", "mammalian_meat_allergy",
               "red_meat_allergy", "meat_allergy")

trends_avg |>
  filter(search_term %in% ags_terms) |>
  ggplot(aes(x = date, y = mean_rsv, colour = search_term_label)) +
  geom_line(linewidth = 0.7) +
  labs(
    title = "AGS/MMA-Specific Search Terms in Australia",
    x = NULL, y = "Mean RSV", colour = "Search Term"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  ) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  scale_colour_brewer(palette = "Set1")

ggsave(file.path(fig_dir, "fig2_ags_terms_comparison.png"),
       width = 25, height = 15, units = "cm", dpi = 300)
```

## 2.3 Summary Statistics per Term

```{r summary-table}
term_summary <- trends_avg |>
  group_by(search_term_label) |>
  summarise(
    first_nonzero = min(date[mean_rsv > 0], na.rm = TRUE),
    peak_month = date[which.max(mean_rsv)],
    peak_rsv = max(mean_rsv, na.rm = TRUE),
    mean_rsv_overall = round(mean(mean_rsv, na.rm = TRUE), 1),
    pct_months_nonzero = round(sum(mean_rsv > 0) / n() * 100, 1),
    .groups = "drop"
  ) |>
  arrange(first_nonzero)

knitr::kable(term_summary, col.names = c(
  "Search Term", "First Non-Zero Month", "Peak Month", "Peak RSV",
  "Mean RSV", "% Months Non-Zero"
), caption = "Summary of search interest by term")
```

## 2.4 Heatmap of Annual Search Interest

```{r fig-heatmap, fig.width=10, fig.height=5}
#| fig-cap: "Annual mean RSV by search term, showing the evolution of AGS awareness over time."

trends_avg |>
  mutate(year = year(date)) |>
  group_by(year, search_term_label) |>
  summarise(annual_mean = mean(mean_rsv, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = year, y = search_term_label, fill = annual_mean)) +
  geom_tile(colour = "white", linewidth = 0.5) +
  scale_fill_viridis_c(option = "inferno", name = "Mean RSV") +
  labs(title = "Annual Search Interest Heatmap", x = "Year", y = NULL) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

ggsave(file.path(fig_dir, "fig3_heatmap.png"),
       width = 25, height = 12, units = "cm", dpi = 300)
```

# 3. Media Coverage Analysis

## 3.1 Monthly Media Coverage Timeline

```{r fig-media-timeline, fig.width=12, fig.height=8}
#| fig-cap: "Monthly media article counts (top) and alpha-gal search interest (bottom), showing temporal relationship between media coverage and public awareness."

# Combine media monthly with alpha_gal search
alpha_gal_monthly <- trends_avg |>
  filter(search_term == "alpha_gal") |>
  dplyr::select(date, mean_rsv)

media_search_combined <- media_monthly |>
  left_join(alpha_gal_monthly, by = "date")

# Faceted panel (avoids dual y-axis interpretability issues)
p_media <- ggplot(media_search_combined, aes(x = date)) +
  geom_col(aes(y = articles), fill = "#d95f02", alpha = 0.7, width = 25) +
  labs(y = "Media Articles", x = NULL) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank()) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y")

p_search <- ggplot(media_search_combined, aes(x = date)) +
  geom_line(aes(y = mean_rsv), colour = "#2c7fb8", linewidth = 0.8) +
  labs(y = "Alpha-gal RSV", x = NULL) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank()) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y")

p_media / p_search +
  plot_annotation(
    title = "Media Coverage vs. Google Search Interest for Alpha-Gal in Australia",
    subtitle = "Top = media article count | Bottom = Google Trends RSV",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(colour = "grey40")
    )
  )

ggsave(file.path(fig_dir, "fig4_media_vs_search.png"),
       width = 28, height = 18, units = "cm", dpi = 300)
```

## 3.2 Key Media Events

```{r media-events-table}
# Filter off-topic articles (xenotransplantation, Lyme disease treatment, French-language)
offtopic_patterns <- regex(
  "pig.*(kidney|transplant)|kidney.*pig|xenotransplant|lyme disease treatment|cypru|les tiques se r|animaux venimeux",
  ignore_case = TRUE
)

media_content_filtered <- media_content |>
  filter(!str_detect(title, offtopic_patterns))

n_excluded <- nrow(media_content) - nrow(media_content_filtered)
cat("Excluded", n_excluded, "off-topic articles (xenotransplantation, Lyme disease",
    "treatment, French-language)\n")
cat("Remaining AGS/MMA-relevant articles:", nrow(media_content_filtered), "\n")

# Rebuild monthly aggregation from filtered content
media_monthly_filtered <- media_content_filtered |>
  mutate(month_date = floor_date(publish_date, "month")) |>
  count(month_date, name = "articles") |>
  rename(date = month_date)

# Use filtered data for all downstream analyses
media_monthly <- media_monthly |>
  dplyr::select(date, total_articles) |>
  left_join(media_monthly_filtered, by = "date") |>
  mutate(articles = replace_na(articles, 0))

# Build media event timeline from filtered content data
media_events <- media_content_filtered |>
  dplyr::select(publish_date, title, media_name) |>
  arrange(publish_date) |>
  distinct(title, .keep_all = TRUE)

knitr::kable(
  media_events |> dplyr::select(publish_date, media_name, title),
  col.names = c("Date", "Source", "Title"),
  caption = "Timeline of Australian media articles on AGS/MMA (unique stories, off-topic excluded)"
)
```

## 3.3 Media Coverage by Year

```{r fig-media-by-year, fig.width=8, fig.height=5}
media_content_filtered |>
  mutate(year = year(publish_date)) |>
  count(year) |>
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = "#d95f02", alpha = 0.8) +
  labs(
    title = "Number of Australian Media Articles on AGS/MMA by Year",
    x = "Year", y = "Article Count"
  ) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

ggsave(file.path(fig_dir, "fig5_media_articles_by_year.png"),
       width = 20, height = 12, units = "cm", dpi = 300)
```

## 3.4 Top Media Sources

```{r}
knitr::kable(
  media_sources |> head(15),
  col.names = c("Source", "Article Count", "Proportion"),
  caption = "Top 15 media sources covering AGS/MMA in Australia",
  digits = 3
)
```

# 4. Seasonality Analysis

## 4.1 Monthly Seasonal Patterns

```{r fig-seasonality, fig.width=12, fig.height=8}
#| fig-cap: "Seasonal patterns by month for tick-related and AGS-specific search terms. Tick season (Oct-Mar, shaded) corresponds to the I. holocyclus activity period."

tick_terms <- c("paralysis_tick", "tick_allergy", "alpha_gal", "mammalian_meat_allergy")

seasonal_data <- trends_avg |>
  filter(search_term %in% tick_terms) |>
  mutate(
    month = month(date, label = TRUE),
    month_num = month(date),
    year = year(date),
    warm_season = ifelse(month_num %in% c(10, 11, 12, 1, 2, 3), "Tick season\n(Oct-Mar)", "Cool season\n(Apr-Sep)")
  )

ggplot(seasonal_data, aes(x = month, y = mean_rsv)) +
  geom_boxplot(aes(fill = warm_season), outlier.size = 1) +
  facet_wrap(~ search_term_label, scales = "free_y", ncol = 2) +
  scale_fill_manual(values = c("Tick season\n(Oct-Mar)" = "#fc8d59",
                                "Cool season\n(Apr-Sep)" = "#91bfdb"),
                    name = "Season") +
  labs(
    title = "Seasonal Distribution of Search Interest",
    subtitle = "Southern Hemisphere tick season: October-March (I. holocyclus activity peak)",
    x = "Month", y = "RSV"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "grey40"),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )

ggsave(file.path(fig_dir, "fig6_seasonality.png"),
       width = 28, height = 20, units = "cm", dpi = 300)
```

## 4.2 Kruskal-Wallis Test for Seasonality

The raw Kruskal-Wallis test can be confounded by secular trends (e.g., months in later years having higher RSV simply because the trend is upward). We therefore also test seasonality after detrending via STL decomposition.

```{r seasonality-tests}
# Raw KW test (may be confounded by trend)
seasonality_raw <- trends_avg |>
  filter(search_term %in% tick_terms) |>
  mutate(month_num = month(date)) |>
  group_by(search_term) |>
  summarise(
    kw_statistic = kruskal.test(mean_rsv ~ factor(month_num))$statistic,
    kw_p_value = kruskal.test(mean_rsv ~ factor(month_num))$p.value,
    .groups = "drop"
  )

# Detrended KW test using STL decomposition
seasonality_detrended <- map_dfr(tick_terms, function(term) {
  term_data <- trends_avg |>
    filter(search_term == term) |>
    arrange(date)

  # Create ts object (monthly, starting from first observation)
  start_year <- year(min(term_data$date))
  start_month <- month(min(term_data$date))
  ts_obj <- ts(term_data$mean_rsv, start = c(start_year, start_month), frequency = 12)

  # STL decomposition to remove trend
  stl_fit <- stl(ts_obj, s.window = "periodic", robust = TRUE)
  detrended <- stl_fit$time.series[, "seasonal"] + stl_fit$time.series[, "remainder"]

  # KW test on detrended values
  month_num <- month(term_data$date)
  kw <- kruskal.test(detrended ~ factor(month_num))

  tibble(
    search_term = term,
    kw_detrended_stat = kw$statistic,
    kw_detrended_p = kw$p.value
  )
})

# Combine raw and detrended results
seasonality_results <- seasonality_raw |>
  left_join(seasonality_detrended, by = "search_term") |>
  mutate(
    sig_raw = ifelse(kw_p_value < 0.05, "Yes", "No"),
    sig_detrended = ifelse(kw_detrended_p < 0.05, "Yes", "No"),
    search_term = term_labels[search_term]
  )

knitr::kable(
  seasonality_results |>
    dplyr::select(search_term, kw_statistic, kw_p_value, sig_raw,
                  kw_detrended_stat, kw_detrended_p, sig_detrended),
  col.names = c("Search Term", "KW Stat (raw)", "p (raw)", "Sig.",
                 "KW Stat (detrended)", "p (detrended)", "Sig."),
  digits = c(0, 2, 4, 0, 2, 4, 0),
  caption = "Kruskal-Wallis test for seasonality: raw vs. STL-detrended RSV"
)
```

## 4.3 Warm vs Cool Season Comparison

```{r season-comparison}
# Wilcoxon rank-sum test for each term
season_test_data <- trends_avg |>
  mutate(
    month_num = month(date),
    warm_season = month_num %in% c(10, 11, 12, 1, 2, 3)
  )

wilcox_results <- season_test_data |>
  group_by(search_term) |>
  summarise(
    wilcox_p = wilcox.test(mean_rsv[warm_season], mean_rsv[!warm_season])$p.value,
    .groups = "drop"
  )

season_comparison <- season_test_data |>
  group_by(search_term, warm_season) |>
  summarise(
    mean = mean(mean_rsv, na.rm = TRUE),
    median = median(mean_rsv, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_wider(
    names_from = warm_season,
    values_from = c(mean, median),
    names_glue = "{.value}_{ifelse(warm_season, 'warm', 'cool')}"
  ) |>
  left_join(wilcox_results, by = "search_term") |>
  mutate(
    warm_cool_ratio = mean_warm / pmax(mean_cool, 0.1),
    search_term = term_labels[search_term]
  ) |>
  arrange(desc(warm_cool_ratio))

knitr::kable(
  season_comparison |>
    dplyr::select(search_term, mean_warm, mean_cool, warm_cool_ratio, wilcox_p),
  col.names = c("Search Term", "Mean RSV (Warm)", "Mean RSV (Cool)",
                 "Warm/Cool Ratio", "Wilcoxon p"),
  digits = c(0, 2, 2, 2, 4),
  caption = "Warm season (Oct-Mar) vs cool season (Apr-Sep) mean RSV with Wilcoxon rank-sum test"
)
```

# 5. Trend Quantification

## 5.1 Quasi-Poisson GAM with Penalised Splines

```{r gam-models}
# Fit GAM models for each search term
gam_results <- list()

for (term in search_terms) {
  term_data <- trends_avg |>
    filter(search_term == term) |>
    mutate(
      time_numeric = as.numeric(date) / 365.25,  # years
      month_num = month(date),
      warm_season = as.integer(month_num %in% c(10, 11, 12, 1, 2, 3)),
      # Add 1 to RSV for quasi-Poisson (avoid log(0))
      rsv_count = round(mean_rsv) + 1
    )

  # Only fit if there's enough non-zero data
  if (sum(term_data$mean_rsv > 0) > 12) {
    gam_fit <- gam(
      rsv_count ~ s(time_numeric, bs = "cr", k = 15) + warm_season,
      family = quasipoisson(),
      data = term_data
    )

    gam_results[[term]] <- list(
      model = gam_fit,
      data = term_data,
      dispersion = summary(gam_fit)$dispersion,
      r_sq = summary(gam_fit)$r.sq,
      warm_effect = coef(gam_fit)["warm_season"],
      warm_p = summary(gam_fit)$p.table["warm_season", "Pr(>|t|)"]
    )
  }
}

# Summary table
gam_summary <- map_dfr(names(gam_results), function(term) {
  r <- gam_results[[term]]
  tibble(
    search_term = term_labels[term],
    r_squared = round(r$r_sq, 3),
    dispersion = round(r$dispersion, 2),
    warm_season_coef = round(r$warm_effect, 3),
    warm_season_irr = round(exp(r$warm_effect), 2),
    warm_season_p = round(r$warm_p, 4)
  )
})

knitr::kable(gam_summary,
             col.names = c("Term", "R-sq", "Dispersion", "Warm Season Coef",
                           "Warm Season IRR", "p-value"),
             caption = "Quasi-Poisson GAM results: trend + warm season effect")

# GAM diagnostics: basis dimension adequacy
cat("\nBasis dimension check (k-index < 1 suggests k may be too low):\n")
gam_diagnostics <- map_dfr(names(gam_results), function(term) {
  r <- gam_results[[term]]
  k_check <- k.check(r$model)
  tibble(
    search_term = term_labels[term],
    k_index = round(k_check[1, "k-index"], 3),
    k_pvalue = round(k_check[1, "p-value"], 4),
    r_sq = round(r$r_sq, 3),
    dispersion = round(r$dispersion, 2),
    fit_quality = case_when(
      r$r_sq < 0.1 ~ "Poor (R-sq < 0.1)",
      r$dispersion > 10 ~ "High overdispersion (>10)",
      TRUE ~ "Adequate"
    )
  )
})

knitr::kable(gam_diagnostics,
             col.names = c("Term", "k-index", "k p-value", "R-sq",
                           "Dispersion", "Fit Quality"),
             caption = "GAM model diagnostics")

cat("\nNote: Models with high overdispersion (>10) may benefit from negative binomial",
    "family.\nModels with R-sq < 0.1 explain negligible variance in the outcome.\n")
```

## 5.2 GAM Trend Visualisation

```{r fig-gam-trends, fig.width=12, fig.height=10}
#| fig-cap: "Quasi-Poisson GAM fitted trends (red) with observed data (blue) for each search term."

gam_plots <- map(names(gam_results), function(term) {
  r <- gam_results[[term]]
  pred_data <- r$data |>
    mutate(fitted = predict(r$model, type = "response") - 1)

  ggplot(pred_data, aes(x = date)) +
    geom_line(aes(y = mean_rsv), colour = "#2c7fb8", alpha = 0.5) +
    geom_line(aes(y = fitted), colour = "#e41a1c", linewidth = 0.8) +
    labs(title = term_labels[term], x = NULL, y = "RSV") +
    theme_minimal(base_size = 9) +
    theme(plot.title = element_text(face = "bold", size = 10))
})

wrap_plots(gam_plots, ncol = 2) +
  plot_annotation(
    title = "Quasi-Poisson GAM Trend Fits",
    subtitle = "Blue = observed mean RSV | Red = GAM fitted trend",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(colour = "grey40")
    )
  )

ggsave(file.path(fig_dir, "fig7_gam_trends.png"),
       width = 30, height = 22, units = "cm", dpi = 300)
```

### Alpha-gal plot for presentation

```{r}
aGal_pres_plot <- gam_plots[[1]] +
  geom_line(aes(y = mean_rsv), colour = MidnightBlue, alpha = 1) +
    geom_line(aes(y = fitted), colour = "darkred", linewidth = 0.8, alpha = 0.5) +
  theme_classic() +
  theme(
    plot.title = element_blank(),
    panel.grid.major.y = element_line(colour = "grey70", size = 0.2),
    panel.background = element_rect(fill = "transparent", colour = "transparent"),
    plot.background = element_rect(fill = "transparent", colour = "transparent")
  )

aGal_pres_plot

ggsave(
  "figures/aGal_pres_plot.png",
  height = 10,
  width = 15,
  units = "cm",
  bg = "transparent",
  dpi = 300
)
```

## 5.3 Segmented Regression (Joinpoint Analysis)

```{r segmented-analysis}
# Aggregate to annual for joinpoint analysis
annual_trends <- trends_avg |>
  mutate(year = year(date)) |>
  group_by(year, search_term) |>
  summarise(annual_mean = mean(mean_rsv, na.rm = TRUE), .groups = "drop")

segmented_results <- list()

for (term in search_terms) {
  term_annual <- annual_trends |>
    filter(search_term == term, annual_mean > 0)

  if (nrow(term_annual) >= 5) {
    lm_fit <- lm(log(annual_mean + 0.1) ~ year, data = term_annual)

    tryCatch({
      seg_fit <- segmented(lm_fit, seg.Z = ~ year, npsi = 1)

      segmented_results[[term]] <- list(
        model = seg_fit,
        breakpoint = round(seg_fit$psi[, "Est."], 1),
        slopes = slope(seg_fit)$year
      )
    }, error = function(e) {
      # No significant joinpoint - use linear model
      segmented_results[[term]] <<- list(
        model = lm_fit,
        breakpoint = NA,
        slopes = data.frame(
          Est. = coef(lm_fit)["year"],
          `Std. Err.` = summary(lm_fit)$coefficients["year", "Std. Error"]
        )
      )
    })
  }
}

# AAPC calculation
aapc_results <- map_dfr(names(segmented_results), function(term) {
  r <- segmented_results[[term]]
  term_annual <- annual_trends |>
    filter(search_term == term, annual_mean > 0)

  if (inherits(r$model, "segmented")) {
    slopes_df <- slope(r$model)$year
    tibble(
      search_term = term_labels[term],
      breakpoint = r$breakpoint,
      slope_before = round((exp(slopes_df[1, "Est."]) - 1) * 100, 1),
      slope_after = round((exp(slopes_df[2, "Est."]) - 1) * 100, 1)
    )
  } else {
    overall_slope <- coef(r$model)["year"]
    tibble(
      search_term = term_labels[term],
      breakpoint = NA_real_,
      slope_before = round((exp(overall_slope) - 1) * 100, 1),
      slope_after = NA_real_
    )
  }
})

knitr::kable(aapc_results,
             col.names = c("Term", "Breakpoint Year", "APC Before (%)",
                           "APC After (%)"),
             caption = "Segmented regression: Annual Percent Change with joinpoint detection")

cat("\nCaveats:\n")
cat("- Segmented regression is fit to annual means (n =",
    n_distinct(annual_trends$year), "years), which limits statistical power.\n")
cat("- Very large APC values (e.g., >100%) for low-volume terms reflect small",
    "absolute changes\n  from near-zero baselines and should be interpreted",
    "as directional rather than precise estimates.\n")
cat("- Terms with median CV >100% (", paste(term_labels[unreliable_terms], collapse = ", "),
    ")\n  have unreliable annual means due to high sampling variability.\n")
```

## 5.4 Joinpoint Visualisation

```{r fig-joinpoint, fig.width=10, fig.height=6}
#| fig-cap: "Annual trends with segmented regression fits for key AGS-related terms."

key_terms <- c("alpha_gal", "mammalian_meat_allergy", "paralysis_tick", "meat_allergy")

annual_plot_data <- annual_trends |>
  filter(search_term %in% key_terms)

ggplot(annual_plot_data, aes(x = year, y = annual_mean)) +
  geom_point(size = 2, colour = "#2c7fb8") +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE,
              colour = "#e41a1c", fill = "#e41a1c", alpha = 0.15) +
  facet_wrap(~ term_labels[search_term], scales = "free_y", ncol = 2) +
  labs(
    title = "Annual Trends in Key Search Terms",
    x = "Year", y = "Annual Mean RSV"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(fig_dir, "fig8_annual_trends.png"),
       width = 25, height = 15, units = "cm", dpi = 300)
```

# 6. Media-Awareness Correlation Analysis

## 6.1 Spearman Correlations

```{r spearman-correlations}
# Merge media monthly with search trends
correlation_data <- trends_avg |>
  dplyr::select(date, search_term, mean_rsv) |>
  pivot_wider(names_from = search_term, values_from = mean_rsv) |>
  left_join(media_monthly |> dplyr::select(date, articles), by = "date")

# Spearman correlation for each term vs media articles
spearman_results <- map_dfr(search_terms, function(term) {
  x <- correlation_data[[term]]
  y <- correlation_data$articles

  # Remove months where both are zero
  valid <- !is.na(x) & !is.na(y)

  test <- cor.test(x[valid], y[valid], method = "spearman", exact = FALSE)

  tibble(
    search_term = term_labels[term],
    rho = round(test$estimate, 3),
    p_value = test$p.value,
    significant = ifelse(test$p.value < 0.05, "Yes", "No")
  )
}) |>
  # Benjamini-Hochberg correction
  mutate(p_adjusted = p.adjust(p_value, method = "BH")) |>
  arrange(desc(abs(rho)))

knitr::kable(spearman_results,
             col.names = c("Search Term", "Spearman rho", "p-value",
                           "Sig. (p<0.05)", "BH-adjusted p"),
             digits = c(0, 3, 4, 0, 4),
             caption = "Spearman correlations between monthly media article count and search RSV (raw levels)")
```

The raw Spearman correlations above may be confounded by shared temporal trends (both media coverage and search interest trend upward). We repeat the analysis on first-differenced series to remove this confound:

```{r spearman-differenced}
# Spearman on first-differenced series (trend-adjusted)
spearman_diff_results <- map_dfr(search_terms, function(term) {
  x <- correlation_data[[term]]
  y <- correlation_data$articles

  valid <- !is.na(x) & !is.na(y)
  x_diff <- diff(x[valid])
  y_diff <- diff(y[valid])

  test <- cor.test(x_diff, y_diff, method = "spearman", exact = FALSE)

  tibble(
    search_term = term_labels[term],
    rho_diff = round(test$estimate, 3),
    p_value_diff = test$p.value,
    significant = ifelse(test$p.value < 0.05, "Yes", "No")
  )
}) |>
  mutate(p_adjusted_diff = p.adjust(p_value_diff, method = "BH")) |>
  arrange(desc(abs(rho_diff)))

knitr::kable(spearman_diff_results,
             col.names = c("Search Term", "Spearman rho (diff)", "p-value",
                           "Sig. (p<0.05)", "BH-adjusted p"),
             digits = c(0, 3, 4, 0, 4),
             caption = "Trend-adjusted Spearman correlations (first-differenced series)")
```

## 6.2 Cross-Correlation Analysis

```{r fig-ccf, fig.width=12, fig.height=8}
#| fig-cap: "Cross-correlation functions between media article counts and search RSV at different monthly lags. Positive lags indicate media leading search interest."

# Focus on key AGS terms
ccf_terms <- c("alpha_gal", "mammalian_meat_allergy", "meat_allergy", "tick_allergy")

ccf_plots <- map(ccf_terms, function(term) {
  term_ts <- correlation_data |>
    dplyr::select(date, all_of(term), articles) |>
    drop_na()

  # Difference both series to ensure stationarity for CCF
  search_diff <- diff(term_ts[[term]])
  media_diff <- diff(term_ts$articles)

  ccf_result <- ccf(media_diff, search_diff, lag.max = 12, plot = FALSE)

  ccf_df <- tibble(
    lag = ccf_result$lag[, 1, 1],
    acf = ccf_result$acf[, 1, 1]
  )

  ci <- qnorm(0.975) / sqrt(length(search_diff))

  ggplot(ccf_df, aes(x = lag, y = acf)) +
    geom_hline(yintercept = 0, colour = "grey50") +
    geom_hline(yintercept = c(-ci, ci), colour = "blue", linetype = "dashed") +
    geom_col(fill = ifelse(abs(ccf_df$acf) > ci, "#e41a1c", "#2c7fb8"),
             width = 0.5) +
    labs(title = term_labels[term], x = "Lag (months)", y = "CCF") +
    theme_minimal(base_size = 9) +
    theme(plot.title = element_text(face = "bold", size = 10))
})

wrap_plots(ccf_plots, ncol = 2) +
  plot_annotation(
    title = "Cross-Correlation: Media Articles vs Search Interest",
    subtitle = "Positive lag = media leads search | Dashed lines = 95% significance bounds",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(colour = "grey40")
    )
  )

ggsave(file.path(fig_dir, "fig9_ccf_analysis.png"),
       width = 28, height = 20, units = "cm", dpi = 300)
```

## 6.3 Optimal Lag Identification

```{r optimal-lag}
optimal_lags <- map_dfr(ccf_terms, function(term) {
  term_ts <- correlation_data |>
    dplyr::select(date, all_of(term), articles) |>
    drop_na()

  search_diff <- diff(term_ts[[term]])
  media_diff <- diff(term_ts$articles)

  ccf_result <- ccf(media_diff, search_diff, lag.max = 6, plot = FALSE)

  ccf_df <- tibble(
    lag = ccf_result$lag[, 1, 1],
    acf = ccf_result$acf[, 1, 1]
  )

  # Find lag with maximum absolute correlation (positive lags only = media leads)
  positive_lags <- ccf_df |> filter(lag >= 0)
  best <- positive_lags |> slice_max(abs(acf), n = 1)

  tibble(
    search_term = term_labels[term],
    optimal_lag = best$lag,
    ccf_at_lag = round(best$acf, 3),
    interpretation = case_when(
      best$lag == 0 ~ "Contemporaneous",
      best$lag > 0 ~ paste("Media leads by", best$lag, "month(s)"),
      TRUE ~ paste("Search leads by", abs(best$lag), "month(s)")
    )
  )
})

knitr::kable(optimal_lags,
             col.names = c("Search Term", "Optimal Lag", "CCF", "Interpretation"),
             caption = "Optimal lag between media coverage and search interest")
```

# 7. Granger Causality Testing

## 7.1 Stationarity Tests

```{r stationarity-tests}
# ADF test (H0: unit root) and KPSS test (H0: stationary) for each series
all_series <- c(ccf_terms, "articles")
series_labels <- c(term_labels[ccf_terms], "articles" = "Media articles")

stationarity_results <- map_dfr(all_series, function(s) {
  ts_data <- correlation_data[[s]] |> na.omit()

  adf <- adf.test(ts_data, alternative = "stationary")
  kpss <- kpss.test(ts_data, null = "Level")

  tibble(
    series = series_labels[s],
    series_id = s,
    adf_statistic = round(adf$statistic, 3),
    adf_p = round(adf$p.value, 4),
    kpss_statistic = round(kpss$statistic, 3),
    kpss_p = round(kpss$p.value, 4),
    adf_stationary = adf$p.value < 0.05,
    kpss_stationary = kpss$p.value >= 0.05,  # fail to reject H0 of stationarity
    conclusion = case_when(
      adf$p.value < 0.05 & kpss$p.value >= 0.05 ~ "Stationary",
      adf$p.value >= 0.05 & kpss$p.value < 0.05 ~ "Non-stationary",
      TRUE ~ "Inconclusive"
    )
  )
})

# Store which series need differencing (for Granger causality)
needs_differencing <- stationarity_results |>
  filter(conclusion != "Stationary") |>
  pull(series_id)

knitr::kable(
  stationarity_results |>
    dplyr::select(series, adf_statistic, adf_p, kpss_statistic, kpss_p, conclusion),
  col.names = c("Series", "ADF Stat", "ADF p", "KPSS Stat", "KPSS p", "Conclusion"),
  digits = c(0, 3, 4, 3, 4, 0),
  caption = "Stationarity tests: ADF (H0: unit root) and KPSS (H0: stationary)"
)
```

## 7.2 VAR Model and Granger Causality

```{r granger-causality}
granger_results <- map_dfr(ccf_terms, function(term) {
  # Create bivariate data frame
  bivar_data <- correlation_data |>
    dplyr::select(date, search = all_of(term), media = articles) |>
    drop_na()

  # Only difference non-stationary series (based on ADF + KPSS results)
  search_vec <- bivar_data$search
  media_vec <- bivar_data$media
  diff_search <- term %in% needs_differencing
  diff_media <- "articles" %in% needs_differencing

  if (diff_search) search_vec <- diff(search_vec)
  if (diff_media) media_vec <- diff(media_vec)

  # Align lengths if only one was differenced
  min_len <- min(length(search_vec), length(media_vec))
  search_vec <- tail(search_vec, min_len)
  media_vec <- tail(media_vec, min_len)

  var_data <- cbind(search_ts = search_vec, media_ts = media_vec)

  tryCatch({
    # Select optimal lag order
    lag_select <- VARselect(var_data, lag.max = 6, type = "const")
    optimal_p <- lag_select$selection["AIC(n)"]
    optimal_p <- max(1, min(optimal_p, 4))  # bound between 1 and 4

    # Fit VAR model
    var_model <- VAR(var_data, p = optimal_p, type = "const")

    # Granger causality tests (both directions)
    gc_media_to_search <- causality(var_model, cause = "media_ts")
    gc_search_to_media <- causality(var_model, cause = "search_ts")

    tibble(
      search_term = term_labels[term],
      differenced = {
        d <- c(if (diff_search) "search", if (diff_media) "media")
        if (length(d) == 0) "none" else paste(d, collapse = "+")
      },
      var_lag = optimal_p,
      media_causes_search_F = round(gc_media_to_search$Granger$statistic, 3),
      media_causes_search_p = gc_media_to_search$Granger$p.value,
      search_causes_media_F = round(gc_search_to_media$Granger$statistic, 3),
      search_causes_media_p = gc_search_to_media$Granger$p.value
    )
  }, error = function(e) {
    tibble(
      search_term = term_labels[term],
      differenced = NA_character_,
      var_lag = NA_integer_,
      media_causes_search_F = NA_real_,
      media_causes_search_p = NA_real_,
      search_causes_media_F = NA_real_,
      search_causes_media_p = NA_real_
    )
  })
})

# Apply BH correction
granger_results <- granger_results |>
  mutate(
    media_to_search_bh = p.adjust(media_causes_search_p, method = "BH"),
    search_to_media_bh = p.adjust(search_causes_media_p, method = "BH")
  )

knitr::kable(
  granger_results |>
    dplyr::select(search_term, differenced, var_lag, media_causes_search_F,
           media_causes_search_p, media_to_search_bh,
           search_causes_media_F, search_causes_media_p, search_to_media_bh),
  col.names = c("Search Term", "Differenced", "VAR Lag", "F (Media->Search)", "p",
                 "BH adj. p", "F (Search->Media)", "p", "BH adj. p"),
  digits = c(0, 0, 0, 2, 4, 4, 2, 4, 4),
  caption = "Granger causality tests: bidirectional media-search relationships (differencing conditioned on stationarity)"
)
```

# 8. Interrupted Time Series Analysis

## 8.1 Identify Key Media Events

```{r media-event-clusters}
# Identify months with media article clusters
media_event_months <- media_monthly |>
  filter(articles > 0) |>
  arrange(desc(articles))

cat("Months with media coverage:\n")
print(media_event_months, n = 20)

# Key events based on content analysis:
# 1. Aug 2023: Major cluster - ticks thriving, meat allergy, US 450K cases
# 2. Nov 2025: Coroner inquest into teen death from MMA
# 3. Jun 2016: Early coverage - thousands allergic to red meat
key_events <- tibble(
  date = as.Date(c("2023-08-01", "2025-11-01", "2018-08-01")),
  event = c(
    "Aug 2023: Tick boom / US 450K cases / SBS & ABC coverage",
    "Nov 2025: NSW coroner inquest into teen MMA death",
    "Aug-Sep 2018: Regional media - tick health risks"
  )
)

knitr::kable(key_events, col.names = c("Date", "Event"),
             caption = "Key identified media events for ITS analysis")
```

## 8.2 ITS Analysis for Aug 2023 Media Event

```{r its-aug2023, fig.width=10, fig.height=6}
#| fig-cap: "Interrupted time series analysis for the August 2023 media event cluster (alpha-gal search term). Red dashed line indicates the intervention point."

# Focus on alpha_gal - the term with most clear trend
its_data <- trends_avg |>
  filter(search_term == "alpha_gal") |>
  mutate(
    time = row_number(),
    post_aug2023 = as.integer(date >= as.Date("2023-08-01")),
    time_after = ifelse(post_aug2023 == 1, time - min(time[post_aug2023 == 1]) + 1, 0),
    month_num = month(date),
    warm_season = as.integer(month_num %in% c(10, 11, 12, 1, 2, 3)),
    rsv_count = round(mean_rsv) + 1
  )

# Segmented quasi-Poisson regression
its_model <- glm(
  rsv_count ~ time + post_aug2023 + time_after + warm_season,
  family = quasipoisson(),
  data = its_data
)

its_summary <- summary(its_model)

# Autocorrelation diagnostic
dw_test <- dwtest(its_model)
cat("Durbin-Watson test for residual autocorrelation: DW =",
    round(dw_test$statistic, 3), ", p =", round(dw_test$p.value, 4), "\n")
if (dw_test$p.value < 0.05) {
  cat("  -> Significant autocorrelation detected. Using Newey-West robust SEs.\n\n")
  nw_vcov <- NeweyWest(its_model, lag = 3, prewhite = FALSE)
  its_coeftest <- coeftest(its_model, vcov. = nw_vcov)
} else {
  cat("  -> No significant autocorrelation. Using model-based SEs.\n\n")
  its_coeftest <- coeftest(its_model)
}

# Report with 95% CIs for IRR
its_params <- c("post_aug2023", "time_after", "warm_season")
its_param_labels <- c("Level change", "Slope change", "Warm season")

cat("ITS Model for Alpha-Gal (Aug 2023 intervention):\n")
cat("=========================================\n")
for (i in seq_along(its_params)) {
  param <- its_params[i]
  est <- its_coeftest[param, "Estimate"]
  se <- its_coeftest[param, "Std. Error"]
  p <- its_coeftest[param, 4]
  irr <- exp(est)
  irr_lower <- exp(est - 1.96 * se)
  irr_upper <- exp(est + 1.96 * se)

  cat(sprintf("  %s: coef = %.3f | IRR = %.2f (95%% CI: %.2f-%.2f) | p = %.4f %s\n",
              its_param_labels[i], est, irr, irr_lower, irr_upper, p,
              ifelse(p < 0.05, "*", "")))
}
cat("Dispersion:", round(its_summary$dispersion, 2), "\n")
n_post_aug2023 <- sum(its_data$post_aug2023)
cat("Post-intervention observations:", n_post_aug2023, "\n")
cat("\nNote: No parameters are statistically significant at alpha = 0.05.",
    "\nThe Aug 2023 media cluster did not produce a detectable step change",
    "in search interest\nbeyond the pre-existing upward trend.\n")

# Visualise
its_data$fitted <- predict(its_model, type = "response") - 1

ggplot(its_data, aes(x = date)) +
  geom_line(aes(y = mean_rsv), colour = "#2c7fb8", alpha = 0.5) +
  geom_point(aes(y = mean_rsv), colour = "#2c7fb8", size = 1, alpha = 0.5) +
  geom_line(aes(y = fitted), colour = "#e41a1c", linewidth = 0.8) +
  geom_vline(xintercept = as.Date("2023-08-01"),
             linetype = "dashed", colour = "red", linewidth = 0.6) +
  annotate("text", x = as.Date("2023-08-01"), y = max(its_data$mean_rsv) * 0.95,
           label = "Aug 2023\nMedia cluster", hjust = -0.1, size = 3,
           colour = "red", fontface = "italic") +
  labs(
    title = "Interrupted Time Series: Alpha-Gal Search Interest",
    subtitle = "Intervention point: August 2023 media coverage cluster",
    x = NULL, y = "RSV",
    caption = "Blue = observed | Red = ITS model fit"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "grey40"),
    plot.caption = element_text(hjust = 0, face = "italic")
  )

ggsave(file.path(fig_dir, "fig10_its_aug2023.png"),
       width = 25, height = 15, units = "cm", dpi = 300)
```

## 8.3 ITS Analysis for Nov 2025 Coroner Inquest

```{r its-nov2025, fig.width=10, fig.height=6}
#| fig-cap: "Interrupted time series analysis for the November 2025 coroner inquest event."

its_data2 <- trends_avg |>
  filter(search_term == "alpha_gal") |>
  mutate(
    time = row_number(),
    post_nov2025 = as.integer(date >= as.Date("2025-11-01")),
    time_after = ifelse(post_nov2025 == 1, time - min(time[post_nov2025 == 1]) + 1, 0),
    month_num = month(date),
    warm_season = as.integer(month_num %in% c(10, 11, 12, 1, 2, 3)),
    rsv_count = round(mean_rsv) + 1
  )

its_model2 <- glm(
  rsv_count ~ time + post_nov2025 + time_after + warm_season,
  family = quasipoisson(),
  data = its_data2
)

its_summary2 <- summary(its_model2)

# Autocorrelation diagnostic
dw_test2 <- dwtest(its_model2)
cat("Durbin-Watson test: DW =", round(dw_test2$statistic, 3),
    ", p =", round(dw_test2$p.value, 4), "\n")
if (dw_test2$p.value < 0.05) {
  cat("  -> Significant autocorrelation. Using Newey-West robust SEs.\n\n")
  nw_vcov2 <- NeweyWest(its_model2, lag = 3, prewhite = FALSE)
  its_coeftest2 <- coeftest(its_model2, vcov. = nw_vcov2)
} else {
  cat("  -> No significant autocorrelation. Using model-based SEs.\n\n")
  its_coeftest2 <- coeftest(its_model2)
}

# Report with 95% CIs
its_params2 <- c("post_nov2025", "time_after", "warm_season")
its_param_labels2 <- c("Level change", "Slope change", "Warm season")

cat("ITS Model for Alpha-Gal (Nov 2025 intervention):\n")
cat("=========================================\n")
for (i in seq_along(its_params2)) {
  param <- its_params2[i]
  est <- its_coeftest2[param, "Estimate"]
  se <- its_coeftest2[param, "Std. Error"]
  p <- its_coeftest2[param, 4]
  irr <- exp(est)
  irr_lower <- exp(est - 1.96 * se)
  irr_upper <- exp(est + 1.96 * se)

  cat(sprintf("  %s: coef = %.3f | IRR = %.2f (95%% CI: %.2f-%.2f) | p = %.4f %s\n",
              its_param_labels2[i], est, irr, irr_lower, irr_upper, p,
              ifelse(p < 0.05, "*", "")))
}
cat("Dispersion:", round(its_summary2$dispersion, 2), "\n")
n_post_nov2025 <- sum(its_data2$post_nov2025)
cat("Post-intervention observations:", n_post_nov2025, "\n")
cat("\nCAUTION: Only", n_post_nov2025, "post-intervention months are available.",
    "\nBernal et al. (2017) recommend a minimum of 8-12 post-intervention",
    "observations\nfor reliable ITS estimation. These results should be",
    "considered preliminary/exploratory\nand will need to be reassessed",
    "as more post-event data accumulate.\n")

its_data2$fitted <- predict(its_model2, type = "response") - 1

ggplot(its_data2, aes(x = date)) +
  geom_line(aes(y = mean_rsv), colour = "#2c7fb8", alpha = 0.5) +
  geom_point(aes(y = mean_rsv), colour = "#2c7fb8", size = 1, alpha = 0.5) +
  geom_line(aes(y = fitted), colour = "#e41a1c", linewidth = 0.8) +
  geom_vline(xintercept = as.Date("2025-11-01"),
             linetype = "dashed", colour = "red", linewidth = 0.6) +
  annotate("text", x = as.Date("2025-11-01"), y = max(its_data2$mean_rsv) * 0.95,
           label = "Nov 2025\nCoroner inquest", hjust = -0.1, size = 3,
           colour = "red", fontface = "italic") +
  labs(
    title = "Interrupted Time Series: Alpha-Gal Search Interest",
    subtitle = "Intervention point: November 2025 NSW coroner inquest into teen MMA death",
    x = NULL, y = "RSV",
    caption = "Blue = observed | Red = ITS model fit"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "grey40"),
    plot.caption = element_text(hjust = 0, face = "italic")
  )

ggsave(file.path(fig_dir, "fig11_its_nov2025.png"),
       width = 25, height = 15, units = "cm", dpi = 300)
```

# 9. Summary of Findings

```{r summary}
# Dynamically extract key findings
ag_peak <- term_summary |> filter(search_term_label == "Alpha-gal")
ag_peak_month <- format(ag_peak$peak_month, "%B %Y")
ag_peak_rsv <- round(ag_peak$peak_rsv, 1)

cat("=" |> strrep(70), "\n")
cat("SUMMARY OF KEY FINDINGS\n")
cat("=" |> strrep(70), "\n\n")

cat("1. SEARCH INTEREST TRENDS\n")
cat("-" |> strrep(40), "\n")
cat("  - Data:", n_distinct(trends_avg$search_term), "search terms, monthly RSV",
    format(min(trends_avg$date), "%Y"), "-", format(max(trends_avg$date), "%Y"),
    "\n    averaged across", max(trends_avg$n_downloads), "daily downloads\n")
cat("  - Alpha-gal search interest peaked at RSV =", ag_peak_rsv, "in", ag_peak_month, "\n")
cat("  - Strong upward trend from ~2020 onwards across AGS-specific terms\n")
cat("  -", n_excluded, "off-topic media articles (xenotransplantation, Lyme)",
    "were excluded\n\n")

cat("2. SAMPLING RELIABILITY\n")
cat("-" |> strrep(40), "\n")
cat("  - Low-volume terms (", paste(term_labels[unreliable_terms], collapse = ", "),
    ")\n    have median CV >100% and results should be interpreted cautiously\n")
cat("  - High-volume terms (food allergy, paralysis tick) have CV <5%\n\n")

cat("3. SEASONALITY\n")
cat("-" |> strrep(40), "\n")
cat("  - Paralysis tick and tick allergy show significant seasonality",
    "(KW p < 0.05)\n    in both raw and detrended analyses\n")
cat("  - Alpha-gal does not show significant seasonality after detrending,\n")
cat("    despite biological plausibility (tick exposure -> awareness)\n\n")

cat("4. MEDIA-AWARENESS RELATIONSHIP\n")
cat("-" |> strrep(40), "\n")
cat("  - Raw Spearman correlations are moderate (rho 0.18-0.32 for AGS terms)\n")
cat("    but may be confounded by shared temporal trends\n")
cat("  - Trend-adjusted (differenced) correlations provide a more conservative estimate\n")
cat("  - Cross-correlation analysis shows contemporaneous (lag 0) association\n\n")

cat("5. GRANGER CAUSALITY\n")
cat("-" |> strrep(40), "\n")
cat("  - No significant Granger causality in either direction after BH correction\n")
cat("  - Differencing was conditioned on stationarity test results (ADF + KPSS)\n\n")

cat("6. KEY MEDIA EVENTS (ITS)\n")
cat("-" |> strrep(40), "\n")
cat("  - Aug 2023 media cluster: NO significant level or slope change (p > 0.05)\n")
cat("    The pre-existing upward trend appears sufficient to explain post-event RSV\n")
cat("  - Nov 2025 coroner inquest: Significant level change (IRR ~2),",
    "but only\n   ", n_post_nov2025, "post-intervention months available",
    "(minimum 8-12 recommended;\n    Bernal et al. 2017). Results are preliminary.\n")
```

```{r session-info}
sessionInfo()
```
